The Two Dualities of Computation: Negative and Fractional Types

### 2012-03-08 rational.tex
_

%subcode{opsem} include main
%! columnStyle = rclr
% <iso, v, C> &|-->& {[iso, v', C]} & (1)
% & & where iso v |--> v' & (2) 
% <c1(;)c2, v, C> &|-->& <c1, v, Fst C c2> & (3) 
% {[c1, v, Fst C c2]} &|-->& <c2, v, Snd c1 C> & (4) 
% {[c2, v, Snd c1 C]} &|-->& {[ c1(;)c2, v, C ]} & (5) 
% <c1(+)c2, left v, C> &|-->& <c1, v, LeftP C c2> & (6) 
% {[ c1, v, LeftP C c2 ]} &|-->& {[c1 (+) c2, left v, C ]} & (7)
% <c1(+)c2, right v, C> &|-->& <c2, v, RightP c1 C> & (8)
% {[ c2, v, RightP c1 C ]} &|-->& {[c1 (+) c2, right v, C ]} & (9)
% <c1(*)c2, (v1, v2), C> &|-->& <c1, v1, LeftT C c2 v2> & (10)
% {[ c1, v1, LeftT C c2 v2 ]} &|-->& <c2, v2, RightT c1 v1 C> & (11) 
% {[ c2, v2, RightT c1 v1 C ]} &|-->& {[ c1 (*) c2, (v1, v2), C ]} & (12)

The rule (1) describes evaluation by a primitive isomorphism. Rules
(2), (3) and (4) deal with sequential evaluation. Rule (2) says that
for the value to flow through the sequence {{c1 (;) c2}} we should
first flow through {{c1}} with {{c2}} pending in the context and rule
(3) says the value that exits from {{c1}} should proceed to flow
through {{c2}}. Rule (4) says that if the value {{v}} has exited
{{c2}} as well, then it is equivalent to having exited from
{{c1(;)c2}}.

Rules (5) to (9) deal with {{c1 (+) c2}} in the same way. In the case
of sums, the shape of the value {{left v}}/{{right v}} determines
whether path {{c1}} or path {{c2}} is taken. Rules (10), (11) and (12)
deal with {{c1 (*) c2}} similarly. In the case of products the value
should have the form {{(v1, v2)}} where {{v1}} flows through {{c1}}
and {{v2}} flows through {{c2}}. Both these paths are entirely
independent of each other and we could chosen to evaluate either
first, or evaluate both in parallel. In this presentation we have
chosen to follow {{c1}] first, but this choice is entirely artificial.

>>>>

%subcode{opsem} include main
%! columnStyle = rclr
% <iso, v, C> &|-->& {[iso, v', C]} & (1)
% & & where iso v |--> v' &
% <c1(;)c2, v, C> &|-->& <c1, v, Fst C c2> & (2) 
% {[c1, v, Fst C c2]} &|-->& <c2, v, Snd c1 C> & (3) 
% {[c2, v, Snd c1 C]} &|-->& {[ c1(;)c2, v, C ]} & (4) 
% <c1(+)c2, left v, C> &|-->& <c1, v, LeftP C c2> & (5) 
% {[ c1, v, LeftP C c2 ]} &|-->& {[c1 (+) c2, left v, C ]} & (6)
% <c1(+)c2, right v, C> &|-->& <c2, v, RightP c1 C> & (7)
% {[ c2, v, RightP c1 C ]} &|-->& {[c1 (+) c2, right v, C ]} & (8)
% <c1(*)c2, (v1, v2), C> &|-->& <c1, v1, LeftT C c2 v2> & (9)
% {[ c1, v1, LeftT C c2 v2 ]} &|-->& <c2, v2, RightT c1 v1 C> & (10) 
% {[ c2, v2, RightT c1 v1 C ]} &|-->& {[ c1 (*) c2, (v1, v2), C ]} & (11)

The rule (1) describes evaluation by a primitive isomorphism. Rules
(2), (3) and (4) deal with sequential evaluation. Rule (2) says that
for the value {{v}} to flow through the sequence {{c1 (;) c2}}, it
should first flow through {{c1}} with {{c2}} pending in the context
({{Fst C c2}}). Rule (3) says the value {{v}} that exits from {{c1}}
should proceed to flow through {{c2}}. Rule (4) says that if the value
{{v}} has exited {{c2}} as well, then it is equivalent to having
exited from {{c1(;)c2}}.

Rules (5) to (8) deal with {{c1 (+) c2}} in the same way. In the case
of sums, the shape of the value {{left v}}/{{right v}} determines
whether path {{c1}} or path {{c2}} is taken. Rules (9), (10) and (11)
deal with {{c1 (*) c2}} similarly. In the case of products the value
should have the form {{(v1, v2)}} where {{v1}} flows through {{c1}}
and {{v2}} flows through {{c2}}. Both these paths are entirely
independent of each other and we could chosen to evaluate either
first, or evaluate both in parallel. In this presentation we have
chosen to follow {{c1}] first, but this choice is entirely artificial.
